Working Title: Predicting Customer Behavior Using a Portuguese Financial Institution Dataset 

### 2024-06-14 code/eda-code/eda - my version.qmd
no more template info

---
title: "An example exploratory analysis script"
date: "2024-02-07"
output: html_document
---

@@

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
```

@@

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata1.rds")
#load data
mydata <- readRDS(data_location)
```

@@

Showing a bit of code to produce and save a summary table.

@@

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

@@

Histogram plots for the continuous outcomes.

Height first.

@@

Now weights.

@@

Now height as function of weight.

@@

Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.

@@

@@

```{r}

mydata |>
  ggplot(aes(job))+
  geom_bar()

@@

mydata |>
  ggplot(aes(marital))+
  geom_bar()

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.

>>>>

---
title: "Exploratory Data Analysis script"
date: "2024-02-07"
output: html_document
author: Antonio Flores
---

@@

# Setup
Necessary libraries for this script.
```{r}
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
```

@@

```{r}
data_location <- here::here("data","processed-data","processeddata1.rds")
mydata <- readRDS(data_location)
```

@@

Creating a summary table.

@@

@@

Histogram plots for some continuous outcomes.

Age first.

@@

Next bank account balance.

@@

Now balance as function of age

@@

Now balance as function of age, stratified by the marital variable.

@@

Next, exploring balance across different job types.

@@

Now, some barplots examining job types, education level, and marital status.

```{r}
mydata |>
  ggplot(aes(job))+
  geom_bar()

@@

mydata |>
  ggplot(aes(marital))+
  geom_bar()
```

### 2024-06-14 code/processing-code/processingfile-v1 - my version.qmd
no more template info

---
title: "An example cleaning script"
author: "Antonio Flores"
date: "2024-13-06"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.

@@

Load needed packages. make sure they are installed.

@@

Note that for functions that come from specific packages (instead of base R), I often specify both package and function like so:
package::function() that's not required one could just call the function specifying the package makes it clearer where the function "lives",
but it adds typing. You can do it either way.

```{r}
# path to data
# note the use of the here() package and not absolute paths

data_location <- here::here("data","raw-data","bank-full.csv")
rawdata <- read.csv(data_location, header = TRUE, sep=";")
```

@@

First we can look at the codebook

@@

Several ways of looking at the data

@@

# Cleaning
```{r}

d1 = rawdata |> 
  mutate(job = as.factor(job)) |>
  mutate(marital = as.factor(marital)) |>
  mutate(education = as.factor(education)) |>
  mutate(default = as.factor(default)) |>
  mutate(housing = as.factor(housing)) |>
  mutate(loan = as.factor(loan)) |>
  mutate(education = as.factor(education)) |>
  mutate(contact = as.factor(contact)) |>
  mutate(month = as.factor(month)) |>
  mutate(poutcome = as.factor(poutcome)) |>
  mutate(y = as.factor(y))

summary(d1)

```

@@

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

@@

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data:
http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

@@

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

>>>>

---
title: "Preparing data for analysis"
author: "Antonio Flores"
date: "2024-13-06"
output: html_document
---

@@

The following packages are needed for this data processing

@@

```{r}
data_location <- here::here("data","raw-data","bank-full.csv")
rawdata <- read.csv(data_location, header = TRUE, sep=";")
```

@@

Examning metadata by accessing the codebook

@@

Various ways of looking at the data

@@

# Cleaning the data

We are mainly converting character variables into factors so we can do some categorical analysis
```{r}
d1 = rawdata |> 
  mutate(job = as.factor(job)) |>
  mutate(marital = as.factor(marital)) |>
  mutate(education = as.factor(education)) |>
  mutate(default = as.factor(default)) |>
  mutate(housing = as.factor(housing)) |>
  mutate(loan = as.factor(loan)) |>
  mutate(education = as.factor(education)) |>
  mutate(contact = as.factor(contact)) |>
  mutate(month = as.factor(month)) |>
  mutate(poutcome = as.factor(poutcome)) |>
  mutate(y = as.factor(y))

summary(d1)
```

@@

Assigning data to output variable.

@@

@@

### 2024-06-14 products/manuscript/Manuscript.qmd
no more template info

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

@@

_To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:_

_Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]._

@@

I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already. Then I will test for correllation between different features and proceed from there.

@@

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

@@

raw = read.csv("C:/Users/Client/Documents/Flores-P2-Project/bank-full.csv",header = TRUE,
                  sep=";")
head(raw)

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Data Snapshot"
#| echo: FALSE
raw=read.csv("../../bank-full.csv", header = TRUE, sep=";")
head(raw)
```

@@

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@@

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

@@

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

@@

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

>>>>

```{r, echo=FALSE, message=FALSE}
library(here)
library(knitr)
```

@@

@@

I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already. Then I will test for correlation between different features and proceed from there.

@@

@@

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Data Snapshot"
#| echo: FALSE
data_location <- here::here("data","raw-data","bank-full.csv")
raw=read.csv(data_location, header = TRUE, sep=";")
head(raw)
```

@@

## Statistical analysis

@@

@@

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable1.rds")
knitr::kable(resulttable)
```

@@

@fig-result shows a scatter plot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Age and bank account balance stratified by marital status"
#| echo: FALSE
knitr::include_graphics(here("results","figures","age-balance-stratified.png"))
```

@@

@@

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
#knitr::kable(resulttable2)
```

@@

## Summary and Interpretation

## Strengths and Limitations

## Conclusions
